Parameter Recovery Validation

### 2025-05-29 nbs/concepts/parameter-recovery-validation/index.qmd
docs: rename validation outputs dir
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

```python
# Create visualization directory
viz_dir = Path("validation_plots")
viz_dir.mkdir(exist_ok=True)

@@

# 1. Parameter recovery diagnostic plots
fig_recovery = plot_parameter_recovery_diagnostics(
    true_parameters=validation_parameters,
    inferred_parameters={k: v["posterior_samples"] for k, v in validation_results.items()},
    figsize=(20, 15)
)
plt.suptitle("Parameter Recovery Diagnostics: All Patterns", fontsize=16)
plt.savefig(viz_dir / "parameter_recovery_diagnostics.png", dpi=300, bbox_inches='tight')
plt.show()

@@

    for dataset_key, adata in pattern_datasets.items():
        fig_phase = plot_piecewise_phase_portrait(
            adata=adata,
            gene_names=adata.var_names[:3].tolist(),  # Plot first 3 genes
            posterior_samples=posterior_samples_all[dataset_key],
            true_parameters=validation_parameters[dataset_key],
            figsize=(15, 10)
        )
        plt.suptitle(f"Phase Portrait: {dataset_key.replace('_', ' ').title()}", fontsize=16)
        plt.savefig(viz_dir / f"phase_portrait_{dataset_key}.png", dpi=300, bbox_inches='tight')
        plt.show()

@@

    for dataset_key, adata in pattern_datasets.items():
        fig_traj = plot_piecewise_trajectories(
            adata=adata,
            gene_names=adata.var_names[:2].tolist(),  # Plot first 2 genes
            posterior_samples=posterior_samples_all[dataset_key],
            time_points=np.linspace(0, 1, 100),
            figsize=(12, 8)
        )
        plt.suptitle(f"Temporal Trajectories: {dataset_key.replace('_', ' ').title()}", fontsize=16)
        plt.savefig(viz_dir / f"trajectories_{dataset_key}.png", dpi=300, bbox_inches='tight')
        plt.show()

# 4. Activation timing analysis
fig_timing = plot_activation_timing(
    posterior_samples=posterior_samples_all,
    true_parameters=validation_parameters,
    figsize=(15, 10)
)
plt.suptitle("Activation Timing Analysis: All Patterns", fontsize=16)
plt.savefig(viz_dir / "activation_timing_analysis.png", dpi=300, bbox_inches='tight')
plt.show()

@@

    for dataset_key, adata in pattern_datasets.items():
        fig_post = plot_posterior_predictive_checks(
            model=model,
            adata=adata,
            posterior_samples=posterior_samples_all[dataset_key],
            figsize=(15, 10)
        )
        plt.suptitle(f"Posterior Predictive Checks: {dataset_key.replace('_', ' ').title()}", fontsize=16)
        plt.savefig(viz_dir / f"posterior_predictive_{dataset_key}.png", dpi=300, bbox_inches='tight')
        plt.show()

print(f"All validation plots saved to {viz_dir}/")
```

>>>>

```python
# Create visualization directory
validation_outputs = Path("validation_plots")
validation_outputs.mkdir(exist_ok=True)

@@

# 1. Parameter recovery diagnostic plots
fig_recovery = plot_parameter_recovery_diagnostics(
    true_parameters=validation_parameters,
    inferred_parameters={k: v["posterior_samples"] for k, v in validation_results.items()},
    figsize=(20, 15)
)
plt.suptitle("Parameter Recovery Diagnostics: All Patterns", fontsize=16)
plt.savefig(validation_outputs / "parameter_recovery_diagnostics.png", dpi=300, bbox_inches='tight')
plt.show()

@@

    for dataset_key, adata in pattern_datasets.items():
        fig_phase = plot_piecewise_phase_portrait(
            adata=adata,
            gene_names=adata.var_names[:3].tolist(),  # Plot first 3 genes
            posterior_samples=posterior_samples_all[dataset_key],
            true_parameters=validation_parameters[dataset_key],
            figsize=(15, 10)
        )
        plt.suptitle(f"Phase Portrait: {dataset_key.replace('_', ' ').title()}", fontsize=16)
        plt.savefig(validation_outputs / f"phase_portrait_{dataset_key}.png", dpi=300, bbox_inches='tight')
        plt.show()

@@

    for dataset_key, adata in pattern_datasets.items():
        fig_traj = plot_piecewise_trajectories(
            adata=adata,
            gene_names=adata.var_names[:2].tolist(),  # Plot first 2 genes
            posterior_samples=posterior_samples_all[dataset_key],
            time_points=np.linspace(0, 1, 100),
            figsize=(12, 8)
        )
        plt.suptitle(f"Temporal Trajectories: {dataset_key.replace('_', ' ').title()}", fontsize=16)
        plt.savefig(validation_outputs / f"trajectories_{dataset_key}.png", dpi=300, bbox_inches='tight')
        plt.show()

# 4. Activation timing analysis
fig_timing = plot_activation_timing(
    posterior_samples=posterior_samples_all,
    true_parameters=validation_parameters,
    figsize=(15, 10)
)
plt.suptitle("Activation Timing Analysis: All Patterns", fontsize=16)
plt.savefig(validation_outputs / "activation_timing_analysis.png", dpi=300, bbox_inches='tight')
plt.show()

@@

    for dataset_key, adata in pattern_datasets.items():
        fig_post = plot_posterior_predictive_checks(
            model=model,
            adata=adata,
            posterior_samples=posterior_samples_all[dataset_key],
            figsize=(15, 10)
        )
        plt.suptitle(f"Posterior Predictive Checks: {dataset_key.replace('_', ' ').title()}", fontsize=16)
        plt.savefig(validation_outputs / f"posterior_predictive_{dataset_key}.png", dpi=300, bbox_inches='tight')
        plt.show()

print(f"All validation plots saved to {validation_outputs}/")
```
